---
title: "Homework2"
author: "Yiheng"
date: "7/14/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## 1.Loading and Cleaning

a.
```{r}
ca_pa <- read.csv("data/calif_penn_2011.csv", header = TRUE)
```
b.
```{r}
nrow(ca_pa)
ncol(ca_pa)
```
$\Rightarrow$ The dataframe has 11275 rows and 34 columns.

c.
```{r}
colSums(apply(ca_pa,c(1,2),is.na))
```
The command shows the number of missing values each column has.

d.
```{r}
ca_pa <- na.omit(ca_pa)
```

e.
```{r}
11275 - nrow(ca_pa)
```
$\Rightarrow$ It eliminated 670 rows.

f. They're compatible, since there might be more than one missing values in one row, and the R script below returns `TRUE`:
```{r}
670 > max(colSums(apply(ca_pa,c(1,2),is.na)))
```

## 2. This Very New House

a.
```{r}
ggplot(data = ca_pa) +
  geom_point(mapping = aes(x = Built_2005_or_later, y = Median_house_value))+
  geom_smooth(mapping = aes(x = Built_2005_or_later, y = Median_house_value))+
  labs(x = "percentage of houses built since 2005",
       y = "median house prices",
       title = "Median House Prices Against New House Percentage") +
  theme_bw()
```

b.
```{r}
ggplot(data = ca_pa) +
  geom_point(mapping = aes(x = Built_2005_or_later, y = Median_house_value))+
  geom_smooth(mapping = aes(x = Built_2005_or_later, y = Median_house_value))+
  labs(x = "percentage of houses built since 2005",
       y = "median house prices",
       title = "Median House Prices Against New House Percentage") +
  facet_wrap(~ STATEFP , scales = "free") +
  theme_bw()
```

## 3. Nobody Home

a.
```{r}
ca_pa <- data.frame(ca_pa, vacancy_rate = ca_pa$Vacant_units/ca_pa$Total_units)
```
```{r}
min(ca_pa$vacancy_rate)
max(ca_pa$vacancy_rate)
mean(ca_pa$vacancy_rate)
median(ca_pa$vacancy_rate)
```
$\Rightarrow$ The minimum vacancy rate is 0;The maximum vacancy rate is 0.965311;  
The mean vacancy rate is 0.08888789;The median vacancy rate is 0.06767283. 

b.
```{r}
ggplot(data = ca_pa) +
  geom_point(mapping = aes(x = Median_house_value, y = vacancy_rate))+
  geom_smooth(mapping = aes(x = Median_house_value, y = vacancy_rate))+
  labs(x = "median house prices",
       y = "vacancy rate",
       title = "Vacancy Rate Against Median House Prices") +
  theme_bw()
```

c.
```{r}
ggplot(data = ca_pa) +
  geom_point(mapping = aes(x = Median_house_value, y = vacancy_rate))+
  geom_smooth(mapping = aes(x = Median_house_value, y = vacancy_rate))+
  labs(x = "median house prices",
       y = "vacancy rate",
       title = "Vacancy Rate Against Median House Prices") +
  facet_wrap(~ STATEFP, scale = "fixed") +
  theme_bw()
```
The plot shows that in California, house vacancy rate and number of houses at house prices between 0 and 1000000\$ are rather even, while in Pennsylvania there are a lot more houses at low prices(0~250000\$) , house vacancy rate is also higher in this range.

## 4.

a.
```{r}
acca <- c()
for (tract in 1:nrow(ca_pa)) {
  if (ca_pa$STATEFP[tract] == 6) {
    if (ca_pa$COUNTYFP[tract] == 1) {
      acca <- c(acca, tract)
    }
  }
}
accamhv <- c()
for (tract in acca) {
  accamhv <- c(accamhv, ca_pa[tract,10])
}
median(accamhv)
```
Variable `acca` in the block of code gives the row indices which the house is in Alameda County (county1 in STATEFP 6), while variable `accamhv` gives the median house values of these houses. Using `median()` function, the whole block gives the median level of house price in Alameda County.

b.
```{r}
median(ca_pa[which(ca_pa$STATEFP == 6 & ca_pa$COUNTYFP ==1),10])
```
c.
```{r}
mean(ca_pa[which((ca_pa$STATEFP == 6 & ca_pa$COUNTYFP ==1)|
              (ca_pa$STATEFP == 6 & ca_pa$COUNTYFP ==85)|
              (ca_pa$STATEFP == 42 & ca_pa$COUNTYFP ==3)),16])
```
$\Rightarrow$ The average percentages of housing built since 2005 is 2.437344.

d.
```{r}
#(i)
cor(ca_pa$Median_house_value,ca_pa$Built_2005_or_later)
#(ii)
cor(ca_pa[which(ca_pa$STATEFP==6),10],ca_pa[which(ca_pa$STATEFP==6),16])
#(iii)
cor(ca_pa[which(ca_pa$STATEFP==42),10],ca_pa[which(ca_pa$STATEFP==42),16])
#(iv)
cor(ca_pa[which(ca_pa$STATEFP == 6 & ca_pa$COUNTYFP == 1),10],
  ca_pa[which(ca_pa$STATEFP==6 & ca_pa$COUNTYFP == 1),16])
#(v)
cor(ca_pa[which(ca_pa$STATEFP == 6 & ca_pa$COUNTYFP == 85),10],
  ca_pa[which(ca_pa$STATEFP==6 & ca_pa$COUNTYFP == 85),16])
#(vi)
cor(ca_pa[which(ca_pa$STATEFP == 42 & ca_pa$COUNTYFP == 3),10],
  ca_pa[which(ca_pa$STATEFP==42 & ca_pa$COUNTYFP == 3),16])
```

e.
```{r}
#extract related information
Alameda <- data.frame(Median_house_value = ca_pa[which(ca_pa$STATEFP == 6 & ca_pa$COUNTYFP == 1),10], Median_income = ca_pa[which(ca_pa$STATEFP == 6 & ca_pa$COUNTYFP == 1),33], county = c("Alameda"))

Santa_Clara <- data.frame(Median_house_value = ca_pa[which(ca_pa$STATEFP == 6 & ca_pa$COUNTYFP == 85),10], Median_income = ca_pa[which(ca_pa$STATEFP == 6 & ca_pa$COUNTYFP == 85),33], county = c("Santa_Clara"))

Allegheny <- data.frame(Median_house_value = ca_pa[which(ca_pa$STATEFP == 42 & ca_pa$COUNTYFP == 3),10], Median_income = ca_pa[which(ca_pa$STATEFP == 42 & ca_pa$COUNTYFP == 3),33], county = c("Allegheny"))

three_counties <- rbind(Alameda, Santa_Clara, Allegheny)

#plot
ggplot(data = three_counties) +
  geom_point(aes(x = Median_income, y = Median_house_value, color = county)) +
  geom_smooth(aes(x = Median_income, y = Median_house_value, color = county)) +
  labs(x = "Median Income",
       y = "Median House Value",
       title = "Median House Value Against Median Income",
       subtitle = "(Chosen counties)") +
  theme_bw()
```

## MB.Ch1.11
Given code block:
```{r}
gender <- factor(c(rep("female", 91), rep("male", 92)))
table(gender)
gender <- factor(gender, levels=c("male", "female"))
table(gender)
gender <- factor(gender, levels=c("Male", "female"))
# Note the mistake: "Male" should be "male"
table(gender)
table(gender, exclude=NULL)
rm(gender)  # Remove gender
```
`table()` function counts the elements by their factor levels, and the factor levels are assumed to be ordered, that's why the first and second output has different orders.

In the third command, `"male"` in the original dataframe `gender` cannot be paired with any of the levels, thus these elements in the output `gender` becomes `NA`:
```{r}
gender <- factor(c(rep("female", 91), rep("male", 92)))
gender <- factor(gender, levels=c("Male", "female"))
is.na(gender[92])
```
In the last command, expression `exclude=NULL` makes `NA` an extra level, and is the last level printed as `<NA>`.

## MB.Ch1.12

```{r}
prop_over <- function(x,c){
  num1 <- 0 
  i <- 1
  while (i <= length(x)){
    if (x[i]>c){
      num1 <- num1+1
    }
    i <- i+1
  }
    prop <- num1/length(x)
  return(prop)
}
```
a.
```{r}
prop_over(seq(1,100),60)
```
b.
```{r}
library(Devore7)
dotplot(ex01.36)
my_ex01.36<- ex01.36$C1
prop_over(my_ex01.36,420)
```

## MB.Ch1.18
```{r}
library(MASS)
```
Using `unstack` function:
```{r}
Value <- unstack(Rabbit, BPchange~Animal)
Treatment <- unstack(Rabbit, Treatment~Animal)[,1]
Dose <- unstack(Rabbit, Dose~Animal)[,1]

Rabbit <- data.frame(Treatment, Dose, Value)
Rabbit
```



