---
title: "Homework3"
author: "Yiheng"
date: "7/14/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


```{r}
gmp <- read.table("data/gmp.dat")
gmp$pop <- round(gmp$gmp/gmp$pcgmp)
```
1.
```{r}
nlmfit1 = 6611*(gmp$gmp/gmp$pcgmp)^(1/8)
nlmfit2 = 6611*(gmp$gmp/gmp$pcgmp)^(0.1)
nlmfit3 = 6611*(gmp$gmp/gmp$pcgmp)^(0.15)
ggplot(data = gmp) +
  geom_point(aes(x = pop, y = pcgmp))+
  labs(x = "Population", y = "Per-Capita GMP ($/person-year)")+
  geom_line(aes(x = pop, y = nlmfit1), col = 'blue', size = 1.5)+
  geom_line(aes(x = pop, y = nlmfit2), col = 'yellow', size = 1.5)+
  geom_line(aes(x = pop, y = nlmfit3), col = 'green', size = 1.5)
```
2.
```{r}
mse <- function(x,N,Y){
  if(missing(N)) N <- gmp$pop
  if(missing(Y)) Y <- gmp$pcgmp
  mse <- sum((Y-x[1]*N^x[2])^2)/length(Y)
  return(mse)
}
mse(c(6611,0.15))
mse(c(5000,0.10))
```
4.
```{r warning=FALSE}
nlm1 <- nlm(mse, c(y0 = 6611, a = 0.15))
nlm2 <- nlm(mse, c(y0 = 6611, a = 1/8))
nlm3 <- nlm(mse, c(y0 = 5000, a = 0.10))

nlm1$minimum
nlm1$estimate
nlm2$minimum
nlm2$estimate
nlm3$minimum
nlm3$estimate
```
`minimum` represents the value of the estimated minimum of `mse()` from the given starting value. `estimate` represents the point at which the minimum value of `mse()` is obtained.

5.
```{r}
plm <- function(x,N,Y){
  nlm_result <- nlm(mse,x)
  plm <- list(nlm_result$estimate[1], nlm_result$estimate[2],nlm_result$minimum)
  names(plm) <- c("final guess for y0","final guss for a", "final value of MSE")
  return(plm)
}
```
Apply `plm()` function to given starting values:
```{r warning=FALSE}
plm(c(6611,0.15))
plm(c(5000,0.10))
```
Two parameter estimates are different because the `plm()` function finds the local optimum.Estimate with starting values `y0 = 6611, a = 0.15` has the lower MSE.

6.a.
```{r}
mean_pcgmp <- mean(gmp$pcgmp)
sem <- sd(gmp$pcgmp)/sqrt(length(gmp$pcgmp))
mean_pcgmp
sem
```
b.
```{r}
except_mean <- function(i){
  return(mean(gmp$pcgmp[-i]))
}
```
c.
```{r}
jackknifed.means <- c()
for (i in 1:length(gmp$pcgmp)){
  jackknifed.means[i] <- except_mean(i)
}
```
d.
```{r}
mean1 <- mean(jackknifed.means)
#jackknife variance
n <- length(gmp$pcgmp)
jackknifed.variance <- (n-1)*sum((jackknifed.means-mean1)^2)/n
#jackknife standard error
jackknifed.sem <- sqrt(jackknifed.variance)
jackknifed.sem
```
Compare jackknifed standard error with `sem` in (a):
```{r}
abs(jackknifed.sem - sem)
```
7.
```{r}
plm.jackknife <- function(x,N = gmp$pop,Y = gmp$pcgmp){
  except_plm <- function(j){
    return(plm(x,N[-j],Y[-j]))
  }
  jackknifed.plm.y0 <- c()
  for (j in 1:length(N)){
    jackknifed.plm.y0[j] <- except_plm(j)[[1]]
  }
  jackknifed.plm.a <- c()
  for (j in 1:length(N)){
    jackknifed.plm.a[j] <- except_plm(j)[[2]]
  }
  mean2 <- mean(jackknifed.plm.y0)
  mean3 <- mean(jackknifed.plm.a)
  #jackknife variance
  m <- length(N)
  jackknifed.variance.plm.y0 <- (m-1)*sum((jackknifed.plm.y0 - mean2)^2)/m
  jackknifed.variance.plm.a <- (m-1)*sum((jackknifed.plm.a - mean3)^2)/m
  result <- c(sqrt(jackknifed.variance.plm.y0),sqrt(jackknifed.variance.plm.a))
  names(result) <- c("se for y0","se for a")
  return(result)
}
```
8.
```{r}
gmp2013 <- read.table("data/gmp-2013.dat")
gmp2013$pop <- round(gmp2013$gmp/gmp2013$pcgmp)
```
Apply `plm()` and `plm.jackknife` functions:
```{r}
plm(c(6611,0.15),gmp2013$pop,gmp2013$pcgmp)
plm.jackknife(c(6611,0.15),gmp2013$pop,gmp2013$pcgmp)
```

