---
title: "Homework1"
author: "Yiheng"
date: "7/13/2020"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## *1.a.*
Make sure current file is in the same folder with folder `data` containing `iowa.csv`. Then use `read.csv` function to load data.
Notice that the default value for `sep` in `read.csv()` function is `","`, claim it as `";"`(which the given csv file uses).
```{r load_data}
iowa.df <- read.csv("data/Iowa.csv", header = TRUE, sep = ";")
```
## *b.* 
Use `nrow()` and `ncol()` function to read the row number and column number:
```{r row_col}
nrow(iowa.df)
ncol(iowa.df)
```
$\Rightarrow$ `iowa.df` has 33 rows and 10 columns.

## *c.*  
Use `colnames()` function to read the names of the columns of `iowa.df`:
```{r}
colnames(iowa.df)
```

## *d.* 
Access element in `iowa.df` with `[]`:
```{r}
iowa.df[5,7]
```

$\Rightarrow$ The value of row 5, column 7 of `iowa.df` is 79.7.

## *e.* 
```{r}
iowa.df[2,]
```

## *2.a.* 
```{r}
vector1 <- c("5", "12", "7", "32")#1
max(vector1)#2
sort(vector1)#3
```

<#1> gives an non-erroneous result, because `c()` function combines values into a *vector* or *list*, which can be in type `character`.

<#2> gives the output `"7"`, since the elements of `vector1` are all of type `character`. The help page of `max()` function says:

>Character versions are sorted lexicographically, and this depends on the collating sequence of the locale in use.  

<#3>`sort()` function sorts a vector into ascending(default) or descending order. For character type elements, they're also sorted lexicographically.

```{r eval=FALSE, message=TRUE}
sum(vector1)#4
```
*Error in sum(vector1) : invalid 'type' (character) of argument*



<#4> should be erroneous, since the arguments of `sum()` function must be numeric or complex or logical vectors, while in this case, it's a character vector.

## *b.*
```{r}
vector2 <- c("5",7,12)#1
```
```{r eval=FALSE, message=TRUE}
vector2[2] + vector2[3] #1
```
*Error in vector2[2] + vector2[3] :* 
  *non-numeric argument to binary operator*

<#1>Error comes from applying binary operator `+` to character type arguments: see the elements of vector2:
```{r}
vector2
```
The output is a vector with elements all of type `character`, since `c()` function coerce all arguments to a common type, which is determined from the highest type of the components in the hierarchy **`NULL`<raw<logical<integer<double<complex<character<list<expression**. In this case, the highest type is `character`.

```{r}
dataframe3 <- data.frame(z1="5",z2=7,z3=12)#2
dataframe3[1,2] + dataframe3[1,3]#2
```
<#2> The result comes from 7+12, since `data.frame` allows columns to have different types, 7 and 12 remain numeric.

```{r}
list4 <- list(z1="6", z2=42, z3="49", z4=126)
list4[[2]]+list4[[4]]#3
```
```{r eval=FALSE, message=TRUE}
list4[2]+list4[4]#4
```
*Error in list4[2] + list4[4] : non-numeric argument to binary operator*


<#3> `list4[[2]]` and `list4[[4]]` are numeric types, as `[[]]` drops names and structures.

<#4> `[]` accesses the elements with names and structures, see output of `list4[2]`:
```{r}
list4[2]
```
It's non-numeric, cannot be applied with binary operator `+`.

## *3.a.*
```{r}
seq(1, 10000, 372)
seq(1, 10000, length = 50)
```

## *b.*
`times` as a single integer, the result consists of the whole input repeat this many times;

`each` as a non-negative integer, each element of `x` is repeated `each` times.

See the difference more directly from output:
```{r}
rep(1:3, times=3)
rep(1:3, each=3)
```
## *MB.Ch1.2*
```{r}
library(DAAG)
library(tidyverse)
```
Use `rbind()` function to extract rows 1,2,4,11,13,18 from `orings`.
```{r}
extracted_dataframe <- rbind(orings[1,],orings[2,],orings[11,],orings[13,],orings[18,],deparse.level = 1)
extracted_dataframe
```
```{r}
ggplot(data = extracted_dataframe) +
  geom_point(aes(x = Temperature,y = Total)) +
  geom_line(aes(x = Temperature, y = Total))+
  labs(y = "Total Incidents",title = "Extracted Total Incidents Against Temperature")+
 theme_bw()
```
```{r}
ggplot(data = orings) +
  geom_point(aes(x = Temperature,y = Total)) +
  geom_line(aes(x = Temperature, y = Total))+
  labs(y = "Total Incidents",title = "Total Incidents Against Temperature")+
 theme_bw()
```

## *MB.Ch1.4.*

## *a.*
```{r}
my_ais <- ais
```
```{r}
str(my_ais)
```
Use `is.na()` function to see if there're any missing values:
```{r}
sum(is.na(my_ais))
```
$\Rightarrow$ None of the columns hold missing values.


## *b.*
```{r}
my_table<- table(my_ais$sex,my_ais$sport)
```
```{r}
names(which((my_table[1,]/my_table[2,])>2))
names(which((my_table[2,]/my_table[1,])>2))
```
$\Rightarrow$ In Gym and Netball the factor of female:male is more than 2:1;
In T_sprnt and W_Polo the factor of male:female is more than 2:1.


## *MB.Ch1.6*

```{r}
a.Manitoba.lakes <- matrix(c(217,254,248,254,253,227,178,207,217,24387,5374,
                             4624,2247,1353,1223,1151,755,657),ncol = 2)
lakenames <-c("Winnipeg","Winnipegosis","Manitoba","SouthernIndian",
              "Cedar","Island","Gods","Cross","Playgreen")#giving row names
rownames(a.Manitoba.lakes)<- lakenames
colnames(a.Manitoba.lakes) <- c("elevation","area")#giving column names
Manitoba.lakes <- data.frame(a.Manitoba.lakes)#transfer into dataframe structure.
Manitoba.lakes
```

## *a.*
```{r}
attach(Manitoba.lakes)
plot(log2(area) ~ elevation, pch=16, xlim=c(170,280))
# NB: Doubling the area increases log2(area) by 1.0
text(log2(area) ~ elevation, labels=row.names(Manitoba.lakes), pos=4)
text(log2(area) ~ elevation, labels=area, pos=2) 
title("Manitoba’s Largest Lakes")
```

## *b.*
```{r}
plot(area ~ elevation, pch=16, xlim=c(170,280), ylog=T)
text(area ~ elevation, labels=row.names(Manitoba.lakes), pos=4, ylog=T)
text(area ~ elevation, labels=area, pos=2, ylog=T) 
title("Manitoba’s Largest Lakes")
```

## *MB.Ch1.7.*

## *a.*
Linear scale:
```{r}
dotchart(Manitoba.lakes$area,
         labels = lakenames,
         main = "Manitoba lakes' area linear dotchart")
```

## *b.*
Logarithmic scale:
```{r}
dotchart(log2(Manitoba.lakes$area),
         labels = lakenames,
         main = "Manitoba lakes' area logarithmic dotchart")
```

## *MB.Ch1.8.*
```{r}
sum(Manitoba.lakes$area)
```
$\Rightarrow$ A lower bound for the area of Manitoba covered by water is 41771 square kilometers.